---
title: "Machine Learning for Prediction: Concepts and Applications in R"
description: |
  Slides, notes, and assignments for FOR TK.
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

```

## Course Title and Instructor

**Title**: FOR TK: Machine Learning for Prediction: Concepts and Applications in R  
**Time and Place**: DAY TIME, LOCATION OR LINK  

**Instructor**: Mike Mahoney  
**Email**: mike.mahoney.218@gmail.com  
**Office Hours**: TK  

## Description

One hour of lecture/discussion per week. High-level introduction to machine 
learning centered on predictive applications of supervised learning, with a 
focus on implementation in R. Topics include: linear and logistic regression, 
tree-based algorithms (inc. CART, random forests, and gradient boosting 
machines), support vector machines, and model averaging.

## Prerequisites

Students must have some familiarity with the R language, including defining 
functions, managing objects, controlling the flow of a program (e.g. if/else 
statements and for-loops), wrangling data in data frames, fitting linear models 
(i.e., the `lm` function) and other basic tasks. 

An introductory stats course (APM391 or equivalent) is recommended.

## Grading

Most weeks, students will be given a hands-on programming assignment due the 
following class (for a total of 11 assignments). Grades are based solely on 
assignment performance; there are no quizzes or exams and no attendance or 
participation requirements. Each assignment is worth an equal number of points 
(approximately 9% of the final grade).

## Schedule

```{r}
library(dplyr)

tribble(
  ~ Week, ~ Topic, ~ Packages, ~ "Homework Assigned",
  "1", "Prediction, Estimation, and Attribution", "None", "Assignment 1",
  "2", "Regression", "None", "Assignment 2",
  "3", "Classification", "pROC", "Assignment 3",
  "4", "Decision Trees", "rpart", "None",
  "5", "Random Forests", "ranger", "Assignment 4",
  "6", "Hyperparameters and Model Tuning", "ranger, purrr", "Assignment 5",
  "7", "Gradient Boosting Machines", "gbm", "Assignment 6",
  "8", "Stochastic GBMs", "lightGBM", "Assignment 7",
  "9", "Loss Functions and Regularization", "lightGBM", "Assignment 8",
  "10", "Support Vector Machines", "kernlab", "Assignment 9",
  "11", "Implementing SVMs", "kernlab", "Assignment 10",
  "12", "Model Averaging", "None", "Assignment 11",
  "13", "Final Topics", "None", "None"
) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

## Textbooks and Materials

This class draws heavily from materials presented in the following two books. 
Both books are freely available online and you do not need to purchase a 
physical copy of either book to succeed in this class; there are no assigned 
readings from these books.

1. Bradley Boehmke and Brandon Greenwell. (2020). Hands on Machine Learning with R. CRC Press. Available online at https://bradleyboehmke.github.io/HOML/ .
2. Bradley Efron and Trevor Hastie. (2016). Computer Age Statistical Inference. Cambridge University Press. Available online at https://web.stanford.edu/~hastie/CASI_files/PDF/casi.pdf .

Students will need to have access to a computer they are able to install R 
packages on.

## Attendance Policy

Attendance is not mandatory for this course. Grades will be assigned based on 
weekly assignments alone; attendance and participation are not incorporated into
grades.

## Students with Learning and Physical Disabilities

SUNY-ESF works with the Office of Disability Services (ODS) at Syracuse University, who is responsible for coordinating disability-related accommodations. Students can contact ODS at 804 University Avenue- Room 309, 315-443-4498 to schedule an appointment and discuss their needs and the process for requesting accommodations. Students may also contact the ESF Office of Student Affairs, 110 Bray Hall, 315-470-6660 for assistance with the process. To learn more about ODS, visit http://disabilityservices.syr.edu. Students who attempt to use accommodations without advance notice to faculty will be referred to the ESF Office of the Dean for Student Affairs. Since accommodations may require early planning and generally are not provided retroactively, please contact ODS as soon as possible.

## Academic Dishonesty

Academic dishonesty is a breach of trust between a student, one’s fellow students, or the instructor(s).  By registering for courses at ESF you acknowledge your awareness of the ESF Code of Student Conduct (http://www.esf.edu/students/handbook/StudentHB.05.pdf ), in particular academic dishonesty includes but is not limited to plagiarism and cheating, and other forms of academic misconduct.  The Academic Integrity Handbook contains further information and guidance (http://www.esf.edu/students/integrity/).  Infractions of the academic integrity code may lead to academic penalties as per the ESF Grading Policy (http://www.esf.edu/provost/policies/documents/GradingPolicy.11.12.2013.pdf).

## Inclusive Excellence Statement

As an institution, we embrace inclusive excellence and the strengths of a diverse and inclusive community. During classroom discussions, we may be challenged by ideas different from our lived experiences and cultures. Understanding individual differences and broader social differences will deepen our understanding of each other and the world around us. In this course, all people  (including but not limited to, people of all races, ethnicities, sexual orientation, gender, gender identity and expression, students undergoing transition, religions, ages, abilities, socioeconomic backgrounds, veteran status, regions and nationalities, intellectual perspectives and political persuasion) are strongly encouraged to respectfully share their unique perspectives and experiences.  This statement is intended to help cultivate a respectful environment, and it should not be used in a way that limits expression or restricts academic freedom at ESF.
