---
title: "Machine Learning for Prediction: Concepts and Applications in R"
description: |
  Slides, notes, and assignments for FOR TK.
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

```

## Course Title and Instructor

**Title**: FOR TK: Machine Learning for Prediction: Concepts and Applications in R  
**Time and Place**: DAY TIME, LOCATION OR LINK  

**Instructor**: Mike Mahoney  
**Email**: mike.mahoney.218@gmail.com  
**Office Hours**: TK  

## Purpose and structure

There exist two cultures in statistical modeling: one looking to explain the 
world, and the other looking to predict it. While statistics -- with an emphasis 
on data models, estimation, and hypothesis testing -- has traditionally focused
on explaining relationships between system inputs and outcomes, the modern world
has given rise to a new class of algorithms which predict outcomes without any 
concept of causality or underlying data models at all.

These pure prediction algorithms, broadly referred to as "machine learning 
algorithms", have been embraced across domains, with massive success and public
interest to match. But this success comes at the cost of complexity, and 
implementing these techniques requires different techniques and a different 
mindset than traditional data model based approaches. 

This course attempts to help with the transition, guiding students through 
several of the most common machine learning approaches with a focus on 
application. Each week, students are expected to read a course handout on the 
concepts and applications of a particular machine learning algorithm. Weekly 
classes will then focus on clarifying and expanding upon the reading, with 
weekly assignments designed to give students hands-on experience implementing 
and tuning these different approaches on both example data and a real-world data 
set of their choice. Topics include the random forest, gradient boosting 
machine, and support vector machine algorithms as well as cross validation, loss 
functions, regularization, and model averaging.

## Prerequisites

Students must have some familiarity with the R language, including defining 
functions, managing objects, controlling the flow of a program (e.g. if/else 
statements and for-loops), wrangling data in data frames, fitting linear models 
(i.e., the `lm` function) and other basic tasks. 

An introductory stats course (APM391 or equivalent) is recommended.

## Calendar

```{r}
library(dplyr)

tribble(
  ~ Week, ~ Topic, ~ Packages, ~ Assignment,
  "1", "Prediction, Estimation, and Attribution", "None", "Assignment 1",
  "2", "Regression", "None", "Assignment 2",
  "3", "Classification", "pROC", "Assignment 3",
  "4", "Decision Trees", "rpart", "None",
  "5", "Random Forests", "ranger", "Assignment 4",
  "6", "Hyperparameters and Model Tuning", "ranger, purrr", "Assignment 5",
  "7", "Gradient Boosting Machines", "gbm", "Assignment 6",
  "8", "Stochastic GBMs", "lightGBM", "Assignment 7",
  "9", "Loss Functions and Regularization", "lightGBM", "Assignment 8",
  "10", "Support Vector Machines", "kernlab", "Assignment 9",
  "11", "Implementing SVMs", "kernlab", "Assignment 10",
  "12", "Model Averaging", "None", "Assignment 11",
  "13", "Final Topics", "None", "None"
) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

## Textbooks

This class draws heavily from materials presented in the following two books. 
Both books are freely available online and you do not need to purchase a 
physical copy of either book to succeed in this class; there are no assigned 
readings from these books.

1. Bradley Boehmke and Brandon Greenwell. (2020). Hands on Machine Learning with R. CRC Press. Available online at https://bradleyboehmke.github.io/HOML/ .
2. Bradley Efron and Trevor Hastie. (2016). Computer Age Statistical Inference. Cambridge University Press. Available online at https://web.stanford.edu/~hastie/CASI_files/PDF/casi.pdf .

## Material Requirements

This course focuses on learning through doing, and as such requires students 
have access to a computer they are able to install software packages on. All
other materials and texts are available for free online.

