---
title: "MLCA Week 9:"
subtitle: "Stochastic Gradient Boosting Machines"  
author: 
  - "Mike Mahoney"
date: "2021-10-27"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(fig.showtext = TRUE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(base_color = "#23395b")
theme_set(theme_xaringan())
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r xaringanExtra-search, echo=FALSE}
xaringanExtra::use_search(show_icon = TRUE)
```

```{r xaringan-tachyons, echo=FALSE}
xaringanExtra::use_tachyons()
```

```{r xaringan-extra-styles, echo = FALSE}
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         #<<
  mute_unhighlighted_code = TRUE  #<<
)
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

class: center, middle

# Stochastic GBMs

---
class: middle

```{r, eval = FALSE}
install.packages("lightgbm")
```

```{r}
library(lightgbm)
```


---
class: middle

# References

---

Titles are links:

+ [HOML](https://bradleyboehmke.github.io/HOML/gbm.html), specifically chapter 
  12 (stochastic GBMs) and 15 (stacked ensembles)
+ [lightGBM parameters](https://sites.google.com/view/lauraepp/parameters)
+ [Model averaging in ecology: a review of Bayesian, information-theoretic, and tactical approaches for predictive inference](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecm.1309)
+ [The Use of Bayesian Model Averaging to Better Represent Uncertainty in Ecological Models](https://www.jstor.org/stable/3588906?seq=1#metadata_info_tab_contents)
