---
title: "terrainr 0.3.0 is out today"
description: |
  New version, who this?
categories: ["R", "Data science", "terrainr"]
author:
  - name: Mike Mahoney
    url: {}
    orcid_id: "0000-0003-2402-304X"
date: 02-17-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    includes: 
      after_body: "../../../utterances.html"
---

[terrainr](https://mikemahoney218.github.io/terrainr/) is 
[in review with rOpenSci](https://github.com/ropensci/software-review/issues/416) 
and the first review just came back! I've been working through the comments over
the past week or so, and today that work has culminated in the release of 
terrainr version 0.3.0.

This is a big release with a handful of breaking changes, so I felt like I 
should give a brief overview of the biggest user-facing changes.

## Breaking Changes

### Object Classes Are Dead; Long Live Object Classes

The single largest change is that 
*terrainr specific classes are no longer exported*, and users shouldn't need to 
worry about getting data into or out of those formats anymore. Instead, use 
any `sf` or `Raster` object in their place. For instance, workflows that used to 
look like this:

```{r eval = FALSE}
# Doesn't run:
library(terrainr)
simulated_data <-  data.frame(id = seq(1, 100, 1),
                              lat = runif(100, 44.04905, 44.17609), 
                              lng = runif(100, -74.01188, -73.83493))

bbox <- get_bbox(lat = simulated_data$lat, lng = simulated_data$lng) 

output_tiles <- get_tiles(bbox = bbox,
                          services = c("elevation", "ortho"),
                          resolution = 90)
```

Now look like this:

```{r}
library(terrainr)
simulated_data <-  data.frame(id = seq(1, 100, 1),
                              lat = runif(100, 44.04905, 44.17609), 
                              lng = runif(100, -74.01188, -73.83493))

simulated_data <- sf::st_as_sf(simulated_data, coords = c("lng", "lat"))
simulated_data <- sf::st_set_crs(simulated_data, 4326)

output_tiles <- get_tiles(data = simulated_data,
                          services = c("elevation", "ortho"),
                          resolution = 90)
```

As part of this change, `get_bbox`, `get_coordinate_bbox`, and all class 
creation and export functions are gone now. Use sf (or Raster*) objects in their
place instead.

### New Names, Who This?

`get_tiles` now uses the `services` argument to name its output list:

```{r}
names(output_tiles)
```

This means that if you request the service `elevation` you can retrieve your 
tiles using the name `elevation`. If you request the same endpoint with 
multiple names, `get_tiles` will use whatever name was first in the vector.

### Fewer Utilities, More Useful

Utility functions `calc_haversine_distance`, `convert_distance`, 
`point_from_distance`, `rad_to_deg`, and `deg_to_rad` have been removed (or 
removed from exports). For unit conversions, check out the 
[units](cran.r-project.org/package=units) package. This shouldn't impact the 
main uses of the package, but is still worth flagging. 

## Show Me What You Got

terrainr 0.3.0 adds a ggplot2 geom, `geom_spatial_rgb`, for plotting 3 band RGB
rasters:

```{r}
library(ggplot2)

ggplot() + 
  geom_spatial_rgb(data = output_tiles[["ortho"]],
                   # Required aesthetics r/g/b specify color bands:
                   aes(x = x, y = y, r = red, g = green, b = blue)) + 
  coord_sf(crs = 4326)
```

Note that above we just passed the file path to our raster; we can also pass a 
RasterStack:

```{r}
ortho <- raster::stack(output_tiles[["ortho"]])

ggplot() + 
  geom_spatial_rgb(data = ortho,
                   aes(x = x, y = y, r = red, g = green, b = blue)) + 
  coord_sf(crs = 4326)
```

Or a data.frame: 

```{r}
ortho_df <- raster::as.data.frame(ortho, xy = TRUE)
names(ortho_df) <- c("x", "y", "red", "green", "blue")

ggplot() + 
  geom_spatial_rgb(data = ortho,
                   aes(x = x, y = y, r = red, g = green, b = blue)) + 
  coord_sf(crs = 4326)
```

Note that each step here gives you a little more control over the output image 
-- for instance, if your raster bands aren't in RGB order (or you have more than
RGBA bands), you'll need to provide a data.frame to get a true color image. 

You can then use these basemaps like most other ggplot geoms:

```{r}
ggplot() + 
  geom_spatial_rgb(data = ortho_df,
                   aes(x = x, y = y, r = red, g = green, b = blue)) + 
  geom_sf(data = simulated_data) + 
  coord_sf(crs = 4326)
```

## New Docs, Who This?

Those are just a few of the changes in 0.3.0; you can find a longer list in the
NEWS file. 

One thing not mentioned in the
[NEWS](https://mikemahoney218.github.io/terrainr/news/index.html) file, though, 
is that this version of terrainr included a complete rewrite of the 
documentation. The docs were mostly written while the package was being 
conceptually developed, and as a result gave a bit too much emphasis to some 
ideas while completely ignoring others. So I've rewritten all of the 
documentation that lives on the 
[terrainr website](https://mikemahoney218.github.io/terrainr/) -- let me know
what you think about the new versions (or if you catch anything I've missed!).
