# Homework Two: Introduction to Data

## Part One: Data Types

### Question 1

Why does this code not work?

```{r}
1 + "a"
```

ANSWER: 

### Question 2

Change the ? in each of these statements so that the statement returns TRUE:

```{r}
4 ? 3

5 ? 6

# Use the not-equal operator here
5 ? 6

# Use the is-equal operator here
5 ? 5

TRUE ? 1

FALSE ? 0

?(NA)
```

### Question 3

Why does this code not return a value? How can we get it to do so? Remember you
can check the help for `sum` by running `?sum`.

```{r}
sum(c(1:10, NA))
```

## Part Two: Subsetting

### Question 3

Use all three of the subsetting methods we've gone over to select the `Species`
column from the `iris` dataset:

```{r}

```

### Question 4

Select the first 6 rows of both the `Species` and `Sepal.Length` column from the 
`iris` dataset. Only one of our subsetting methods will work for this, but you 
can specify the columns two different ways -- do it both ways in the chunk below!

```{r}

```

### Question 5

Create a data frame and a list with the same data: a single vector named x, 
containing the numbers 1 through 10. Save them (using `<-`) in our_data_frame 
and our_list, respectively.

What's the difference between the output of `head` when called on a list versus
a data frame?

```{r}

```

ANSWER:

### Question 6

Use our subsetting techniques to pull the number 6 out of `our_list`. Do it 
two times, once using `$` and once only using `[[]]`.

```{r}

```

