---
title: 2013 NYC Flights Dashboard
author: Mike Mahoney
params:
  flights: NA
  graph_flights_per_hour: NA
  graph_mean_delay: NA
  graph_sankey: NA
  graph_carriers: NA
output:
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    orientation: rows
---
# EWR {data-navmenu="Airport Overviews"}

##

###

```{r warning=FALSE}
airportcode <- "EWR"

graph_flights_per_hour(airportcode)
```

###
```{r warning=FALSE}
graph_sankey(airportcode)
```

###
```{r warning=FALSE}
graph_carriers(airportcode)
```

## {.tabset}

### `r month.name[[1]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 1)
```


### `r month.name[[2]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 2)
```


### `r month.name[[3]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 3)
```


### `r month.name[[4]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 4)
```


### `r month.name[[5]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 5)
```


### `r month.name[[6]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 6)
```


### `r month.name[[7]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 7)
```


### `r month.name[[8]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 8)
```


### `r month.name[[9]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 9)
```


### `r month.name[[10]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 10)
```


### `r month.name[[11]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 11)
```


### `r month.name[[12]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 12)
```

# LGA {data-navmenu="Airport Overviews"}

##

###

```{r warning=FALSE}
airportcode <- "LGA"

graph_flights_per_hour(airportcode)
```

###
```{r warning=FALSE}
graph_sankey(airportcode)
```

###
```{r warning=FALSE}
graph_carriers(airportcode)
```

## {.tabset}

### `r month.name[[1]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 1)
```


### `r month.name[[2]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 2)
```


### `r month.name[[3]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 3)
```


### `r month.name[[4]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 4)
```


### `r month.name[[5]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 5)
```


### `r month.name[[6]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 6)
```


### `r month.name[[7]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 7)
```


### `r month.name[[8]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 8)
```


### `r month.name[[9]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 9)
```


### `r month.name[[10]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 10)
```


### `r month.name[[11]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 11)
```


### `r month.name[[12]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 12)
```

# JFK {data-navmenu="Airport Overviews"}

##

###

```{r warning=FALSE}
airportcode <- "JFK"

graph_flights_per_hour(airportcode)
```

###
```{r warning=FALSE}
graph_sankey(airportcode)
```

###
```{r warning=FALSE}
graph_carriers(airportcode)
```

## {.tabset}

### `r month.name[[1]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 1)
```


### `r month.name[[2]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 2)
```


### `r month.name[[3]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 3)
```


### `r month.name[[4]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 4)
```


### `r month.name[[5]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 5)
```


### `r month.name[[6]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 6)
```


### `r month.name[[7]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 7)
```


### `r month.name[[8]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 8)
```


### `r month.name[[9]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 9)
```


### `r month.name[[10]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 10)
```


### `r month.name[[11]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 11)
```


### `r month.name[[12]]`

```{r, fig.width=16, fig.height=3.5, warning=FALSE}
graph_mean_delay(airportcode, 12)
```

