---
title: "A semi-comprehensive `automap` walkthrough"
output: html_document
---

This vignette aims to cover the various arguments to `automap`, the main 
function within the `spacey` package.

This vignette will only discuss functions which aren't comprehensively covered 
in the [package introduction](https://mikemahoney218.github.io/spacey/articles/introduction-to-spacey.html).
If you're stuck on saving and loading data or generating a map of the right 
place at all, check that document out first.

## Map Generation and Sizing

### Image Sizing

There are two arguments to `automap` which control the output size of your image,
somewhat predictably named `img.width` (for the x dimension) and `img.height` 
(for the y). Both of these measurements are in pixels and have maximum values of
8000 (enforced by the API, not just this package).

### Overlay

An overlay is a png image which will be placed over your elevation map, 
downloaded from the ESRI MapServer API. The following overlays can be accessed:

* "World_Imagery"
* "NatGeo_World_Map"
* "USA_Topo_Maps"
* "World_Physical_Map"
* "World_Shaded_Relief"
* "World_Street_Map"
* "World_Terrain_Base"
* "World_Topo_Map"

You can see previews of many of these overlays at 
[this link](http://leaflet-extras.github.io/leaflet-providers/preview/index.html)
by looking for the maps prefixed with "ESRI". Note that not all ESRI maps on 
that list are available through the public API used by this package.

The transparency of the overlay layer is controlled via `overlay.alpha`, with a
default value of 0.75. A value of 1 means the overlay will be fully opaque 
(looking like Google Maps), while a value of 0 will be fully transparent.

### Elevation Scale

The z-scale (units of your x/y data divided by units of your z data) for USGS 
data is approximately 9, the default value used by the package. This will result
in your map's elevations being roughly accurately scaled, which can sometimes 
make for lame maps. To exaggerate elevation features, try setting `z` to lower 
values (I've gone as low as 0.0001 for artistic renderings before); 
alternatively, to flatten your map set `z` to higher values.

### Displaying the Map

By default, `automap` prints your map and then invisibly returns the map itself.
To suppress the printing, set `print.map = FALSE` in your call to `automap`.

## Map Coloring

This section will make comparsions against a simple map to show the effect each 
argument has on coloration. This is the map we're using as our default:
```{r}
library(spacey)
automap(
  44.121268,
  -73.903734
)
```

Shadow intensity can be controlled via the `max.darken` argument. Lower values 
will result in higher contrast images:

```{r}
automap(44.121268,
  -73.903734,
  max.darken = 0.1
)
```

The height of the sun used to draw shadows can be changed via the `sun.altitude`
function:
```{r}
automap(44.121268,
  -73.903734,
  sun.altitude = 90
)
```

The direction of the sun, meanwhile, can be changed via `sun.angle`. Values here
represent degree values -- so 0 means the sun is located due North, 90 due East,
and so on.
```{r}
automap(44.121268,
  -73.903734,
  sun.angle = 115
)
```

Shading colors can be controlled via the `colorscale` argument.
All individual scales implemented in `rayshader` (at the time of this 
writing, this meant `imhof1` through `imhof4`, `desert`, `bw`, and `unicorn`) 
can be provided as values to `colorscale`.  
Additionally, this package implements three new scales for maps:

```{r}
automap(44.121268,
  -73.903734,
  colorscale = "spacey1"
)
```

```{r}
automap(44.121268,
  -73.903734,
  colorscale = "spacey2"
)
```

```{r}
automap(44.121268,
  -73.903734,
  colorscale = "spacey3"
)
```

I'm taking suggestions for better names for these color schemes (though the 
numeric ones will be left as aliases) -- suggestions are welcome on 
[GitHub](https://github.com/mikemahoney218/spacey/issues)!

## 3D Options

I personally find it easiest to render 3D images in interactive sessions, then
use the interactive viewer to position my map as I want it. I'll then use
`rayshader::render_snapshot()` to save the image from my interactive viewer to 
file. Better support for scripted 3D rendering is expected in future package 
versions.
